### 2015-02-18 Paper.tex
Add templates

\documentclass[11pt, ngerman, fleqn, DIV=15, headinclude]{scrartcl}

\usepackage[bibatend, color]{header}

\hypersetup{
    pdftitle=
}

%\subject{}
\title{}
%\subtitle{}
\author{
    Martin Ueding \\ \small{\href{mailto:mu@martin-ueding.de}{mu@martin-ueding.de}}
}

\begin{document}

\maketitle

%\newpage
%\tableofcontents
%\newpage

\end{document}

% vim: spell spelllang=de tw=79