### 2015-03-13 Paper.tex
Work on paper and presentation

\documentclass[11pt, english, fleqn, DIV=15, headinclude]{scrartcl}

\usepackage[bibatend, color]{header}

\usepackage{pdflscape}
\usepackage[section]{placeins}

\hypersetup{
    pdftitle=
}

\newcommand\mpi{m_{\piup}}
\newcommand\mpipi{m_{\piup\piup}}

\subject{physics760 Computational Physics}
\title{Analysis of $\piup$-$\piup$ lattice QCD scattering data}
%\subtitle{}
\author{
    Martin Ueding \\ \small{\href{mailto:mu@martin-ueding.de}{mu@martin-ueding.de}}
}

\begin{document}

\maketitle

\begin{abstract}
    % TODO
\end{abstract}

\tableofcontents

\newpage

\section{Data generation}

\subsection{Metropolis algorithm}

\subsection{Correlation functions}

\section{Analysis methods}

Figure~\ref{fig:analysis-flow} shows the data flow in the analysis. This
section will go through the whole analysis in the order of the flow chart. The
methods used will be explained along the way when they are needed.

\begin{figure}[htbp]
    \centering
    \includegraphics[width=\linewidth]{sketches/Zeichnung.pdf}
    \caption{%
        Data flow in the analysis. The first step is the import of the data,
        which is covered in Section~\ref{sec:import}. The two and four point
        functions are combined into pairs for each configuration. The data is
        then folded in half. The bootstrap step is covered in
        Section~\ref{sec:bootstrap}. Then a correlated fit is performed, see
        Section~\ref{sec:correlated_fit} for details. The second to last step
        is the application of Lüscher's formula, see
        Section~\ref{sec:scattering_length}. The results are shown in
        Section~\ref{sec:results}.
    }
    \label{fig:analysis-flow}
\end{figure}

\subsection{Import data}
\label{sec:import}

The data that I was given is organized in different ensembles (like A30.32,
A100.24). For each ensemble, multiple configurations were simulated. The
output of each configuration is one two-point correlation function $C_\piup(t)$
and three four-point correlation functions $C_{\piup\piup}^{(i)}(t)$. Those
three different contractions are combined into a single
$\piup\piup$-correlation function:
\begin{equation}
    C_{\piup\piup}(t) = C_{\piup\piup}^{(1)}(t) + C_{\piup\piup}^{(2)}(t)
    - 2 C_{\piup\piup}^{(3)}(t).
\end{equation}

Some of the ensembles had multiple versions of the data in it. I analyzed all
of them sorted by file name. In the tables, you will find the ensembles
multiple times, those are the different versions.

The number of configurations in each ensemble is called $N$.

In both space and time, the lattice used in the computation has periodic
boundary conditions. The spatial lattice extent is called $L$, the temporal one
$T$. This leads to a symmetry where the time slices $t$ and $T-t$ closely
related. The correlation functions $C(t)$ with $t = |t_1 - t_2|$ are symmetric
in $t_1$ and $t_2$, such that only the data in the interval $[0, T/2]$ carries
independent information. Before starting any other calculations with the data,
it is folded in half and averaged. The data then looks like in
Figure~\ref{fig:folded} which is shown later on.

\subsection{Bootstrap}
\label{sec:bootstrap}

All error estimation is done with the bootstrap method. One other known method
to compute errors is the Gaussian error propagation. This method involves
computing derivatives of the functions applied to the data, which can be either
numerically unstable or just not really applicable: How does the variance of
the mean change when you change the input data?

In order to get rid of all those problems at the same time one does a meta
analysis of the data itself. To get back to the example of the mean, let $X$ be
the set of $N$ data points ($x_i \in \R$). Then the estimator of the mean
$\bar\mu$ can be computed directly from $X$. Now $R$ bootstrap samples $X^r$
are computed from the set by randomly selecting $N$ elements from $X$ and
putting them into $X^r$ (upper index, no power). Since $X$ is sampled from the
original distribution, it resembles the original distribution as good as it can
with only $N$ elements. $X^r$ therefore is sampled from an approximation of the
original distribution. Then the estimators for the means $\bar\mu^r$ are
computed from the bootstrapped sets. The mean of all the $\bar\mu^r$ should
give $\bar\mu$, which was computed from the original data. The error estimate
is then the standard deviation of all the $\bar\mu^r$.

The number of bootstrap samples $R$ was set to $R = 3N$ for each ensemble.

\subsection{Correlated fit}
\label{sec:correlated_fit}

The pion masses are contained in the correlation functions in the form of an
exponential decay constant. Since time is periodic in this simulation, the
correlation function is also periodic with the temporal lattice extent $T$.
Therefore, the expectation is not a simple exponential decay—$\exp(-\lambda
t)$—but a cosh-like function. The four-point functions also have a constant
contribution due to the finite lattice extent. For both two and four point
functions, the functions that I fitted to the folded data with parameters $\vec
\lambda$ are:
\begin{equation}
    \label{eq:fit2}
    m_\piup(t, \lambda) = \lambda_1 \sbr{\exp(-\lambda_2 t) + \exp(-\lambda_2
    [T - t])}
\end{equation}
and
\begin{equation}
    \label{eq:fit4}
    m_{\piup\piup}(t, \lambda) = \lambda_1 \sbr{\exp(-\lambda_2 t) + \exp(-\lambda_2
    [T - t])} + \lambda_3.
\end{equation}

The folded correlation functions with the fit functions are shown in
Figure~\ref{fig:folded}.

\begin{figure}[htbp]
    \centering
    \begin{minipage}[c]{0.49\linewidth}
        \centering
        \includegraphics[width=\linewidth]{plots/A100_24_L24_T48_beta190_mul0100_musig150_mudel190_kappa1632550__ev120__TB2_SO_LI6_new_c2_folded.pdf}
    \end{minipage}
    \hfill
    \begin{minipage}[c]{0.49\linewidth}
        \centering
        \includegraphics[width=\linewidth]{plots/A100_24_L24_T48_beta190_mul0100_musig150_mudel190_kappa1632550__ev120__TB2_SO_LI6_new_c4_folded.pdf}
    \end{minipage}
    \caption{%
        Folded correlation functions with cosh-like fit function using a
        logarithmic ordinate scale. On the left side, the two point function is
        shown, the four point function on the right. Only the blue data points
        were used for fitting. The right ordinate shows the residuals in great
        magnification (red). One can see that the errors seem way too large.
        This is a result of the autocorrelation of the data points.
    }
    \label{fig:folded}
\end{figure}

The correlation functions themselves are highly correlated with respect to
time. Therefore, a regular least squared fit would give a $\chi^2$ that would
be way too low. The $p$-values usually end up around 1, which does not imply a
perfect fit but rather that the data fits the model too \emph{well}. This, in
turn, means that errors and residuals are over-estimated with the given model.
A different model is called for: the correlated fit.

For the correlated fit, a new likelihood function is needed. I chose to keep
the least squared likelihood function but incorporate the correlation into the
$\chi^2$. First, a correlation matrix $\tens C$ is needed, which is computed
from the $R$ bootstrap samples:
\begin{equation}
    \label{eq:correlation_matrix}
    C_{ij} := \frac{1}{R[R-1]} \sum_{k=1}^R
    [x_{ik} - \bar x_{iR}] [x_{jk} - \bar x_{jR}]
    \eqnsep
    \bar x_{iR} := \frac 1R \sum_{r=1}^R x_{ir}.
\end{equation}

Using this correlation matrix, a new $\chi^2$ can be defined which incorporates
the inverse correlation matrix:
\begin{equation}
    \label{eq:chisq}
    \chi^2_\text{corr} := \sum_{i, j}^T
    \left[ \bar x_{iR} - f(t_i, \lambda) \right]
    C^{-1}_{ij}
    \left[ \bar x_{jR} - f(t_j, \lambda) \right].
\end{equation}
The regular $\chi^2$ has $\tens C^{-1} = \tens 1$ and is just the sum of the
squared residuals.

In the curve fitting process, a function like
\texttt{scipy.optimize.curve\_fit} tries to find the parameters $\lambda$ such
that $\chi^2$ becomes minimal. In my implementation, I used
\texttt{scipy.optimize.leastsq} which tries to minimize the squared norm of a
vector valued function. With a Cholesky decomposition, this can be done like
so: Define the square bracket in Eq.~\eqref{eq:chisq} to be the residual vector
$\vec r$. Then the matrix multiplication can be written as
\begin{equation}
    \chi^2 = \vec r^\text T \tens C^{-1} \vec r.
\end{equation}
The Cholesky decomposition gives me $\tens C^{-1} = \tens U^\dagger \tens U$,
where $\tens U$ is a upper triangle matrix. Then I can write
\begin{equation}
    \chi^2 = [\tens U \vec r]^\dagger [\tens U \vec r]
    = \| \tens U \vec r \|^2
\end{equation}
where $\tens U \vec r$ is a vector that can be fed into
\texttt{scipy.optimize.leastsq}.

As shown in Figure~\ref{fig:analysis-flow}, the correlation matrix computed from
the bootstrap samples is also used to fit the original data. Each fit yields a
value for $\mpi$ and $\mpipi$. The computed masses are shown in
Table~\ref{tab:masses}.

\subsection{Scattering length}
\label{sec:scattering_length}

There is a relation between mass difference to scattering length
\parencite[(1.3)]{luescher/volume_dependence}:
\begin{equation}
    \label{eq:luescher}
    \mpipi = 2\mpi - \frac{4 \piup a_0}{\mpi L^3} \sbr{1 + c_1 \frac{a_0}L + c_2 \frac{a_0^2}{L^2}}
    \eqnsep
    c_1 = \num{-2.837297}
    \eqnsep
    c_2 = \num{6.375183}
\end{equation}

Using the computed $\mpi$ and $\mpipi$ with Lüscher's formula, I can
numerically solve for the scattering length $a_0$. In my program, I use
\texttt{scipy.optimize.brentq} that is based on the algorithm by Brent (1973).

The computed scattering lengths are shown in Table~\ref{tab:masses}.

\section{Results}
\label{sec:results}

\begin{table}
    \centering
    \begin{tabular}{lSSS}
        ensemble & {$m_{\piup}$}  & {$m_{\piup\piup}$} & {$a_0$} \\
        \midrule
        A100.24 & 0.22238 +- 0.00023 & 0.45125 +- 0.00052 & -1.346 +- 0.029 \\
        A100.24 & 0.22233 +- 0.00040 & 0.45083 +- 0.00095 & -1.287 +- 0.132 \\
        A100.24 & 0.22239 +- 0.00024 & 0.45111 +- 0.00053 & -1.316 +- 0.039 \\
        A30.32  & 0.12416 +- 0.00055 & 0.25135 +- 0.00144 & -0.904 +- 0.273 \\
        A40.20  & 0.14779 +- 0.00078 & 0.31429 +- 0.00162 & -1.426 +- 0.052 \\
        A40.24  & 0.14447 +- 0.00052 & 0.29815 +- 0.00109 & -1.255 +- 0.051 \\
        A40.24  & 0.14453 +- 0.00031 & 0.29842 +- 0.00071 & -1.272 +- 0.051 \\
        A40.32  & 0.14126 +- 0.00022 & 0.28626 +- 0.00055 & -1.228 +- 0.092 \\
        A60.24  & 0.17275 +- 0.00052 & 0.35278 +- 0.00126 & -1.194 +- 0.112 \\
        A60.24  & 0.17279 +- 0.00048 & 0.35404 +- 0.00099 & -1.361 +- 0.091 \\
        A80.24  & 0.19930 +- 0.00024 & 0.40511 +- 0.00057 & -1.228 +- 0.043 \\
        B55.32  & 0.15553 +- 0.00022 & 0.31589 +- 0.00055 & -1.676 +- 0.117 \\
        D45.32  & 0.12047 +- 0.00046 & 0.25057 +- 0.00137 & -2.416 +- 0.205
    \end{tabular}
    \caption{%
        Computed masses from correlation functions. The last column shows the
        scattering length $a_0$ which is computed using Lüscher's formula,
        Eq.~\eqref{eq:luescher}.
    }
    \label{tab:masses}
\end{table}

\begin{table}
    \centering
    \begin{tabular}{lSSSS}
        ensemble & {$L$} & {$T$} & {$a_0 m_\piup$} & {$m_\piup/f_\piup$} \\
        \midrule
        A100.24 & 24 & 48 & -0.2993 +- 0.0066 & 2.77 \\
        A100.24 & 24 & 48 & -0.2862 +- 0.0292 & 2.77 \\
        A100.24 & 24 & 48 & -0.2927 +- 0.0088 & 2.77 \\
        A30.32  & 32 & 64 & -0.1122 +- 0.0339 & 1.86 \\
        A40.20  & 20 & 48 & -0.2107 +- 0.0077 & 2.11 \\
        A40.24  & 24 & 48 & -0.1813 +- 0.0074 & 2.03 \\
        A40.24  & 24 & 48 & -0.1839 +- 0.0074 & 2.03 \\
        A40.32  & 32 & 64 & -0.1735 +- 0.0131 & 2.06 \\
        A60.24  & 24 & 48 & -0.2063 +- 0.0194 & 2.32 \\
        A60.24  & 24 & 48 & -0.2351 +- 0.0157 & 2.32 \\
        A80.24  & 24 & 48 & -0.2448 +- 0.0087 & 2.55 \\
        B55.32  & 32 & 64 & -0.2607 +- 0.0182 & 2.34 \\
        D45.32  & 32 & 64 & -0.2911 +- 0.0248 & 2.49
    \end{tabular}
    \caption{%
        Lattice size of the ensembles together with computed quantities.
        These data points are also shown in Figure~\ref{fig:result}.
        The pion
        decay constants are taken from
        \parencite[table~1]{Knippschild/Pi_Pi_Scattering}.
    }
    \label{tab:computed}
\end{table}

The results from Table~\ref{tab:computed} are shown in Figure~\ref{fig:result}.

\begin{figure}[htbp]
    \centering
    \includegraphics[width=\linewidth]{plots/result.pdf}
    \caption{%
        $\mpi a_0$ as a function of $\mpi/f_\piup$ as shown in
        Table~\ref{tab:computed}. Diamond data points are taken from draft
        paper from Carsten Urbach's workgroup. Some ensembles had multiple
        versions of the correlation functions in them, this plot shows all of
        them in lexical order of the pathname.
    }
    \label{fig:result}
\end{figure}

\end{document}

% vim: spell spelllang=en tw=79